Energy aware HPC

### 2021-07-08 thesis/intro/intro.tex
update thesis text

\section{Motivation}
The energy efficiency of data centers has attained key importance in recent years, due to its high economic, environmental, and performance impact. 
Data center consumption was estimated in 2010 at between 1.1\% and 1.5\% of the worldwide electricity usage \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation like Argentina  \cite{Mathew2012Energy-awareNetworks}.
Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
Nowadays, the leading petaflop supercomputers consume a range of 1–18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today’s data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}. 
Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.

@@

The rest of this paper is divided as follows. In Section \ref{sec:relatedwork}, a general review of models is provided, showing the difference between each approach and its applications. In Section \ref{sec:models}, the models and its parameters are derived alongside its constraints. In section \ref{sec:experimentalvalidation}, the model is validated with the PARSEC benchmark applications. Forward, in Section \ref{sec:dvfs_optmin}, use cases of the model are presented, how it could be applied in DVFS and also analysis of the best saving energy policy depending on the system and the effect of the parallelism in the energy. Finally, the conclusion is make in Section \ref{sec:conclusion}.

>>>>

\section{Motivation}
The energy efficiency of data centers has attained key importance in recent years, due to its high economic, environmental, and performance impact. 
Data center consumption was estimated in 2010 at between 1.1\% and 1.5\% of the worldwide electricity usage \cite{Dayarathna2016DataSurvey,Corcoran2017EmergingICT}, generating as much pollution as a nation like Argentina  \cite{Mathew2012Energy-awareNetworks}.
Power costs exceed in some cases the cost of purchasing hardware \cite{Rivoire2007ModelsOptimizations}. 
Nowadays, the leading petaflop supercomputers consume a range of 1-18 MW of electrical power, with 1.5 MW on average, which can be easily translated into millions of dollars per year in electricity bills \cite{Group2012HandbookSahni}.
Furthermore, the energy costs of powering a typical data center doubles every five years \cite{Buyya2013MasteringProgramming}.
Therefore, with such steep increase in electricity use and rising electricity costs, electricity bills have become a major expense for today's data centers \cite{Poess2008EnergyResults}, \cite{Gao2013QualityCenters}. 
Due to these reasons, data center energy efficiency is now considered a chief concern for data center operators, often ahead of the traditional considerations of availability and security.

@@

%The rest of this paper is divided as follows. In Section \ref{sec:relatedwork}, a general review of models is provided, showing the difference between each approach and its applications. In Section \ref{sec:models}, the models and its parameters are derived alongside its constraints. In section \ref{sec:experimentalvalidation}, the model is validated with the PARSEC benchmark applications. Forward, in Section \ref{sec:dvfs_optmin}, use cases of the model are presented, how it could be applied in DVFS and also analysis of the best saving energy policy depending on the system and the effect of the parallelism in the energy. Finally, the conclusion is make in Section \ref{sec:conclusion}.

### 2021-07-08 thesis/main.tex
first version
add phases in main
update thesis text

% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{nth}
\usepackage{todonotes}
\usepackage{url}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{float}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{listings}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage[export]{adjustbox}

@@

\usepackage{suthesis-2e}
\dept{Computer Science}

\begin{document}
\title{Energy aware HPC}
\author{Vitor Ramos Gomes da Silva}
\principaladviser{Carlos Valderrama}
\coprincipaladviser{Pierre MANNEBACK}
\firstreader{Thierry DUTOIT}
\secondreader{Sidi MAHMOUDI}
\thirdreader{Samuel Xavier de Souza} %if needed

\beforepreface
\prefacesection{Preface}
Second Phd Committee
%\prefacesection{Acknowledgments}
%I would like to thank...
\afterpreface

\chapter{Introduction}
Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation. This work proposes an analytical model that can be used to estimate energy-optimal software operation. 
This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
\include{intro/intro}

@@

\chapter{Models}
In this Section, the proposed power, performance and energy models are derived and validated
\include{models/models}
\include{ch5/ch5}

\chapter{Optmizers}
This chapter talks about the different types of optimizers.
\include{ch7/ch7}

>>>>

% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{nth}
\usepackage{todonotes}
\usepackage{url}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{float}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{listings}

%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage[export]{adjustbox}

\lstset { %
	language=C++,
	backgroundcolor=\color{black!5}, % set backgroundcolor
	basicstyle=\footnotesize,% basic font setting
}

@@

%\usepackage{suthesis-2e}
%\dept{Computer Science}

\usepackage[fpms]{umons-coverpage}
\umonsAuthor{Vitor Ramos Gomes \textsc{SILVA}}
\umonsTitle{Energy aware high-performance computing}
\umonsDocumentType{A DISSERTATION SUBMITTED TO THE ELECTRONICS AND MICROELECTRONICS DEPARTMENT OF MONS UNIVERSITY IN PARTIAL FULFILLMENT OF THE REQUIREMENTS FOR THE DEGREE OF DOCTOR IN ENGINEERING SCIENCES AND TECHNOLOGY}
%\umonsSupervisor{Supervisor\\ Carlos Alberto \textsc{VALDERRAMA}\\
%				 Co-supervisor\\ Pierre \textsc{MANEBACK}}
\umonsSupervisor{
 		Thesis committee:\\
 		Supervisor: Prof. Carlos Alberto \textsc{VALDERRAMA}\\
 		Co-supervisor: Prof. Pierre \textsc{MANEBACK}\\
 		Prof. Thierry \textsc{DUTOIT}\\
 		Prof. Sidi \textsc{MAHMOUDI}\\
 		Prof. Samuel \textsc{XAVIER-DE-SOUZA}
}
\umonsDate{April 2021}

\begin{document}
%\title{Energy aware HPC}
%\author{Vitor Ramos Gomes da Silva}
%\principaladviser{Carlos Valderrama}
%\coprincipaladviser{Pierre MANNEBACK}
%\firstreader{Thierry DUTOIT}
%\secondreader{Sidi MAHMOUDI}
%\thirdreader{Samuel Xavier de Souza} %if needed
% 
%\beforepreface
%\prefacesection{Preface}
%%Second Phd Committee
%\prefacesection{Acknowledgments}
%%I would like to thank...
%\afterpreface

\umonsCoverPage

%\section*{Acknowledgments}
%\vspace{1cm}
%\begin{center}
%	...
%\end{center}
%\newpage

\section*{Abstract}
Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation. This work proposes an analytical model that can be used to estimate energy-optimal software operation. 
This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.

\vspace{2cm}
\textbf{Keywords :} Energy Model; Dynamic Frequency and Voltage Scaling; Dynamic Power Management; High Performance Computing
\newpage

\tableofcontents
\newpage
\listoffigures
\newpage
\listoftables
\newpage

\chapter{Introduction}

\include{intro/intro}

@@

\chapter{Models}
In this Section, the proposed power, performance and energy models are derived and validated
\include{models/models}
\include{model_analysis/model_analysis}

\chapter{Optmizers}
This chapter talks about the different types of optimizers.
\include{optmizers/optmizers}
\include{phases/phases}

### 2021-07-08 thesis/phases/phases.tex
add phases in main
update thesis text

\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
\caption{Energy and number of cores with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
    \label{fig:cores_control}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
\caption{Energy and frequency with 60 and 70 phases. (a) bla bla (b) bla bla (c) blabla.}
    \label{fig:freq_control}
\end{figure}

>>>>

\begin{figure}[h!]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_cores_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_cores_signals_cmp.png}
%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_cores_signals_cmp.png}  
\caption{Energy and number of cores with 37 and 99 phases.}
    \label{fig:cores_control}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_black_3_freq_signals_cmp.png}
\includegraphics[width=\columnwidth]{phases/figures/signals/completo_bodytrack_3_freq_signals_cmp.png}
%\includegraphics[width=\columnwidth]{phases/figures/signals/completo_canneal_4_freq_signals_cmp.png}
\caption{Energy and frequency with 37 and 99 phases.}
    \label{fig:freq_control}
\end{figure}
